---
title: "488 Lab 3"
author: "Fred Kaesmann"
date: "March 24, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE,
               message = FALSE,
               warning = FALSE,
               include = FALSE)
```

```{r names}
##### Names Vector ####
# This will be used multiple times below
namesVec <- 
    c("Water (gals)", 
      "Malted Barley (lbs)", 
      "Hops (lbs)", 
      "Brewer's Yeast (pkgs)")
```


\clearpage

Ray,

We looked over your production arrangements as requested and we are reasonably
certain we have found some significant improvements to your current production
model.

From the outset we would like you to know that we understand the realities 
of your current situation, resource management has 
become a top priority in the industry given the recent shuttering of bars 
and restaurants throughout the tri state area.
Other breweries we have worked with have seen their revenue take a 40% to 50%
hit, forcing them to furlough employees and cut pay for those remaining.
Then on top of it are the difficult discussions with investors. 

We believe that given your existing output, a small batch brewery like yourself
could be well positioned to withstand a market downturn, assuming you are 
willing to make some tough decisions.

### The Numbers

By the information you provided we have your current production at 
100 gallons of Ale, 400 gallons of Lager and 200 gallons of Stout, from which
your current monthly profit is \$17,734.
This can be improved dramatically, though we must note this will come at the 
cost of some difficult business decisions.

Our proprietary model has developed an improved production plan relative to the 
monthly profit by house beer.
\footnote{Though we need to stress that our model does not take into account the 
time required for brewing, nor the labor required for each beer, as the 
information was not provided. 
We would be more than happy to have our model account for 
this if you would like to provide that information.}
The implementation of this plan will require some changes not only to production
but will also require changes at the brand level, so should you decide to
pursue this to fruition, we highly recommend including a representative from 
marketing and sales to discuss the impact on the brand image, in addition to
the bottom line. 

Our model has developed a production plan which will, by our calculations, 
improve your montly profit by approximately 20% to \$21,507.
We stress again that this is not a "free lunch," though this represents a 
marked improvements in the bottom line, this assumes you cease production 
of your Copy Cat "We Swear it's not a Repackaged Guinness" stout, on more than 
just copyright infringement grounds. 
We again stress the need to include marketing and sales in on this decision
before moving forward, as our model has not looked into market sentiment
surrounding each of your products.
\footnote{We feel now is the appropriate time to mention that this is another
one of the many forms of analysis we can provide upon request.}

Our model maximizes your profit with 337 gallons of your "Haughty Nurse Ale" and 
524 gallons of your "Sam Clemens Style Lager."

### The New Deal

We also looked into your questions about the profitablility of entering into a 
strategic partnership with the Two Dog Brewery down the road. 
Our model shows this being a very green new deal, as this will increase 
your monthly profits, on top of what the model has already shown above. 
Should you enter into the agreement with Two Dog we find that would increase
profits an additional 10% to \$23463.


As always it was a pleasure working with you gentlemen, and if you have any
further questions feel free to contact us



Fred Kaesmann

President and CEO 

Brewing by the Numbers 

\clearpage
# Appendix I
## Methods and Calculations

We approached this as a simple exercise in linear programming.

Given the initial revenue function
\begin{equation*}
  \text{Monthly Revenue } = \$38 \text{ per gallon of Ale} + \$26 
  \text{ per gallon of Lager} + \$40.50 \text{ per gallon of Stout}
\end{equation*}

the ingredient requirements 
```{r ingredient_matrix, include=TRUE}
##### Create Ingredient Matrix ####
ingredientMatrix <- matrix(
             c(1.250, 1.100, 1.300,
               2.250, 1.222, 3.233,
               0.045, 0.011, 0.022,
               0.157, 0.185, 0.172),
             byrow = TRUE,
  nrow = 4,
  dimnames = list(
    # Row Names
    namesVec, 
    # Column Names
    c("Ale", "Lager", "Stout"))
)

kable(ingredientMatrix)
```

as well as the associated costs
```{r cost_vector, include=TRUE}
##### Cost Vector ####
costVector <- matrix(
             c(0.45,
               2.50,
               3.75, 
               6.00),
             nrow = 4,
             byrow = TRUE,
             dimnames = list(
               namesVec,
               "Material Costs"
             ))
kable(costVector)
```

we were able to combine the cost vector and the ingredient matrix to develop a 
cost matrix.

```{r cost_matrix, include=TRUE}
##### Matrix of Costs ####
costMatrix <- ingredientMatrix * costVector[, 1]
kable(costMatrix)
```

This cost matrix yields our cost per style of beer,

\begin{equation*}
  \text{Monthly Cost } = \$7.30 \text{ per gallon of Ale} +  \$4.70  \text{ per gallon of Lager} + 
  \$9.78 \text{ per gallon of Stout}
\end{equation*}

which, when subtracted from the above revenue equation, gives us the
montly profit for the brewery.

\begin{equation*}
  \text{Monthly Profit } = \$30.70 \text{ per gallon of Ale} + \$21.30 
  \text{ per gallon of Lager} + \$30.72 \text{ per gallon of Stout}
\end{equation*}

```{r profit}
##### Profit ####
# Objective function
# Profit = Revenue - Cost
profit <- c(38 - sum(costMatrix[, 1]), 
            26 - sum(costMatrix[, 2]),
         40.50 - sum(costMatrix[, 3]))


##### Constraint Directionals ####
directionals <- c("<=", 
                  "<=", 
                  "<=", 
                  "<=")
```

\clearpage
Our given system constrains we were,
```{r avail_vec, include=TRUE}
##### Availablility Vector ####
availablilityVector <- matrix(c(
    2000, 
    1400, 
    50, 
    150),
    byrow = TRUE,
    dimnames = list(
      namesVec,
      "Availability"
    )
)

# Nice display
kable(availablilityVector)
```

and when combined with the cost per resource gives us the budgetary impact of 
our resource constraints. 

```{r cost_constraint, include=TRUE}
##### Spending Limits ####
costConstraint <- availablilityVector * costVector

colnames(costConstraint) <- "Availablility (Dollars)"
kable(costConstraint)
```

Taking all these factors into account, our model finds that profit is maximized 
at \$21,507 when Copy Cat makes 337 gallons of Ale, 524 gallons of Lager.
Our model also returns shadow prices of \$4.16 for Malted Barley and \$7.28 
for yeast packages, meaning for each additional pound of Malted Barley and 
each additional yeast package the brewery can expect to see increases in monthly
profit by \$4.16 and \$7.28 respectively. 
```{r solve}
##### Solve with lp() ####
solutions <- lpSolve::lp("max",
                profit,
                costMatrix,
                directionals,
                costConstraint,
                int.vec = c(1:3),
                compute.sens = TRUE)

##### Explore Solutions Object ####
cat("Our x,y,z values are:\n", solutions$solution)

cat("\nThe profit is:\n", solutions$objval)
    
cat("\nThe Shadow Prices are:\n", solutions$duals)
```

### The New Deal
After the deal offered by Two Dog, we readjust our resource availability, 
recognizing the increased availability of Malted Barley and additional yeast
packages.
```{r new_availability, include=TRUE}
##### New Availability Vector ####
new_availability <- matrix(
  c(
    2000, 
    2000, 
    50, 
    155),
  nrow = 4,
  dimnames = list(
    namesVec,
    "New Availability"
  )
)
# Nice display
kable(new_availability)
```

We solve as before using the new resource constraints, and find that even 
after the new deal, we see increases in profits by an additional 10% to \$23463.
```{r new_solutions}
##### New Solutions ####
new_solutions <- lpSolve::lp("max",
                profit,
                ingredientMatrix,
                directionals,
                new_availability,
                compute.sens = TRUE)

##### Explore Solutions Object ####
cat("Our producction levels of Ale, Lagger, and Stout (repsectively) are:\n", 
    new_solutions$solution)

cat("\nThe Profit after the deal is:\n", 
    new_solutions$objval - 600 * 7.5)
    
    
cat("\nThe Shadow Prices are:\n", 
    new_solutions$duals)
```
We are left with new shadow prices for Malted Barley and Yeast Packages, 
\$10.41 and \$46.37 respectively. 


\clearpage
# Appendix II

```{r appendix_code, echo=TRUE, eval=FALSE, include=TRUE}

##### Chunk Options ####
library(knitr)                  # chunk options and kable() for fancy display
opts_chunk$set(echo = TRUE,
               message = FALSE,
               warning = FALSE)

##### Names Vector ####
# This will be used multiple times below
namesVec <- 
    c("Water (gals)", 
      "Malted Barley (lbs)", 
      "Hops (lbs)", 
      "Brewer's Yeast (pkgs)")

##### Create Ingredient Matrix ####
ingredientMatrix <- matrix(
             c(1.250, 1.100, 1.300,
               2.250, 1.222, 3.233,
               0.045, 0.011, 0.022,
               0.157, 0.185, 0.172),
             byrow = TRUE,
  nrow = 4,
  dimnames = list(
    # Row Names
    namesVec, 
    # Column Names
    c("Ale", "Lager", "Stout"))
)
kable(ingredientMatrix)


##### Cost Vector ####
costVector <- matrix(
             c(0.45,
               2.50,
               3.75, 
               6.00),
             nrow = 4,
             byrow = TRUE,
             dimnames = list(
               namesVec,
               "Material Costs"
             ))
kable(costVector)

costMatrix <- ingredientMatrix * costVector[, 1]
kable(costMatrix)

##### Profit ####
# Objective function
# Profit = Revenue - Cost
profit <- c(38 - sum(costMatrix[, 1]), # Ale 
            26 - sum(costMatrix[, 2]), # Lager
         40.50 - sum(costMatrix[, 3])) # Stout

##### Constraint Directionals ####
directionals <- c("<=", 
                  "<=", 
                  "<=", 
                  "<=")

##### Availablility Vector ####
availablilityVector <- matrix(c(
    2000, 
    1400, 
    50, 
    150),
    byrow = TRUE,
    dimnames = list(
      namesVec,
      "Availability"
    )
)
# Nice display
kable(availablilityVector)

##### Spending Limits ####
costConstraint <- availablilityVector * costVector
colnames(costConstraint) <- "Availablility (Dollars)"
kable(costConstraint)

##### Solve with lp() ####
solutions <- lpSolve::lp("max",
                profit,
                costMatrix,
                directionals,
                costConstraint,
                compute.sens = TRUE)

##### Explore Solutions Object ####
cat("Our x,y,z values are:\n", solutions$solution)

cat("\nThe profit is:\n", solutions$objval)
    
cat("\nThe Shadow Prices are:\n", solutions$duals)

##### New Availability Vector ####
new_availability <- matrix(
  c(
    2000, 
    2000, 
    50, 
    155),
  nrow = 4,
  dimnames = list(
    namesVec,
    "New Availability"
  )
)
# Nice display
kable(new_availability)

##### New Solutions ####
new_solutions <- lpSolve::lp("max",
                profit,
                ingredientMatrix,
                directionals,
                new_availability,
                compute.sens = TRUE)

##### Explore Solutions Object ####
cat("Our producction levels of Ale, Lagger, and Stout (repsectively) are:\n", 
    new_solutions$solution)

cat("\nThe Profit after the deal is:\n", 
    new_solutions$objval - 600 * 7.5)
    
    
cat("\nThe Shadow Prices are:\n", 
    new_solutions$duals)

```
