---
title: "MAT 488 Project 5"
author: "Fred Kaesmann"
date: "May 2020"
output: pdf_document
---

\begin{center}
  \textit{
    But it is only for the sake of profit that any man employs capital in the support of industry; and he will always, therefore, endeavour to employ it in the support of that industry of which the produce is likely to be of the greatest value, or to echange for the greatest quantity either of money or of other goods. 
  } \\ - Adam Smith
\end{center}

```{r setup, include=FALSE}
##### Chunk Options ####
library(knitr)
opts_chunk$set(
     echo = FALSE,
  include = FALSE,
  message = FALSE,
  warning = FALSE,
  # Figure Settings 
  fig.asp = 0.618,
fig.width = 6)
```

```{r packages}
##### Packages ####
library(FredsStatsPack)  # Our Very Own
library(magrittr)
library(data.table)
library(ggplot2)
```

\section{The Problem}

Due to the recent economic downturn, Wile E. Cyote's Agricultural supplier of 
choice, Acme Tractors, has been forced to limit their capital
expenditures.
To address this problem they have contacted us at Fred's 
Executive Research and Statistical Analysis: $t$ to $z,$ to assist with 
inventory optimization. 

As the trade war tensions have heated, the possibility of increases in domestic 
agriculture has risen dramatically.
Acme is trying to better prepare themselves by taking full advantage
of their obviously meager resources, and looking to perform an external review
of their inventory management practices. 
Extensive evaluation by Acme's internal analysts say they have on average about 
one tractor sale per week. 
Given their current facilities they are only capable of housing up to five
tractors at any given time, and tractor shipments only come in sets of five 
tractors, so they cannot schedule a new delivery until they are completely 
out of units.
Deliverys can only be made on Saturdays and Sundays, so if they run out of 
tractors they must wait until the weekend before their supply is replenished.

\section{Markov Modeling Supply and Demand}
\begin{center}
  \textit{
    And by directing that industry in such a manner as its produce may be of 
    the greatest value, he intends only his own gain, and he is in this, as in 
    many other cases, led by an invisible hand to promote an end which was no
    part of his intention. 
  }
\end{center}

Acme can house a discrete number of tractors and at the end of the week can 
only be left with a discrete number of tractors. 
Given the assumption of a Poisson distribution, and an assumed mean value, 
$\lambda_{\text{Poisson}},$
\footnote{Not to be confused with $\lambda_{\text{Sensitivity}}$}
 we know the probability of some event occuring.
In the context of this problem, we can model the demand for tractors by the 
equation 
\begin{equation*}
  P(X = k) = \frac{\lambda^k e^{-\lambda}}{k!}
\end{equation*}



As stated previously, their internal analysts have determined they sell about
one tractor per week. 
We can adjust our above distribution and model Acme's tractor demand by

\begin{equation*}
  P(X = k) = \frac{e^{-1}}{k!}
\end{equation*}

\subsection{Markov Chains}
We have a small set of possible initial states and discrete probabilities 
associated with transitions between states. 
This makes Markov Chains a preferred method for evaluating the possiblilities.

```{r initial_conditions, include=TRUE}
##### Initial Conditions ####
# Poisson Probability
# We know lambda = 1
lambda <- 1
p <- function(k){lambda^k * exp(-lambda)/factorial(k)}

# Transition matrix
transition_matrix <- matrix(
  c(p(0),    0,    0,    0, 1 - p(0),
    p(1), p(0),    0,    0, 1 - sum(sapply(c(0:1), p)),
    p(2), p(1), p(0),    0, 1 - sum(sapply(c(0:2), p)),
    p(3), p(2), p(1), p(0), 1 - sum(sapply(c(0:3), p)),
    p(4), p(3), p(2), p(1), 1 - sum(sapply(c(1:4), p))
    ),
  byrow = TRUE,
  nrow = 5,
  ncol = 5)

# We take as the initial condition the last row
#   of the transition matrix
initial_state <- transition_matrix[5, ]
```

```{r transition_table, include=TRUE}
# Display transition matrix
kable(round(transition_matrix * 100, 3), caption = "Transition Matrix")
```

Table 2 contains our projection for the first three weeks and the tenth week
after beginning fully stocked with five tractors. 
We find that by the third week they are most likely to have 5 tractors in stock.
```{r markov_chain, include=TRUE}
##### Markov Chain Matrix ####
state_mat <- matrix(nrow = 11, ncol = 5,
                    dimnames = list(
                      c("Initial Week", paste("Week", c(1:10))),
                      c("1 Tractor", paste(c(2:5), "Tractors"))))

# Start with initial condition 
# Fill matrix with weekly position
for (week in c(0:10)) {
  state_mat[week+1, ] <- Freds_MarkovChain(initial_state, transition_matrix, week)
}

# x-axis for our plot
week <- data.table(c(0:10))
colnames(week) <- "Week"

##### Inventory Projection Table ####
kable(round(state_mat[c(1:4, 11), ] * 100, 3), caption = "Inventory Probability Matrix")

```

The chart below plots the probability of $n$ tractors in stock by week. 
Again we see the probability of 5 tractors (the purple line) peaking by week
5 then leveling out and approaching our steady state. 

```{r inventory_plot, include=TRUE}
##### Plot Markov Chain ####
state_mat %>%
  cbind(week) %>%
  as.data.table() %>%
  melt.data.table(id.vars = "Week") %>%
  ggplot(aes(x = Week,
             y = value * 100,
             color = variable)) +
  geom_line() +
  xlab("Week") +
  ylab("Probability of k Tractors Week Start") + 
  ggtitle("Tractor Inventory Projections")
```


\subsection{Steady State Analysis}
We evaluate our model to determine if our tractors inventory likelihood
converges to a steady state. 
This reflects the position we expect to arrive at after any variation from
our initial position has been "washed out" by repeated application of our 
transition matrix.
The table below represents our calculated convergence for tractor inventory.
\footnote{The following were evaluated to a tolerance of $10e^{-17}$ just because.}

```{r steady_state, include=TRUE}
##### Determine Steady State ####
# Find the steady state for tolerance 10e-17
steady_state <- Freds_MarkovChain_SteadyState(initial_state, transition_matrix, -17, 100)

steady_state <- matrix(steady_state,
                       nrow = 1,
                       dimnames = list(
                         "",
                         c("1 Tractor", paste(c(2:5), "Tractors"))
                       ))

##### Format Steady State nicely #####
kable(round(steady_state * 100, 3), caption = "Tractor Steady State")
```

Our steady state largely reflects our visualization of our weekly projections.
We see here again that we are most likely to have 5 tractors in inventory and 
least likely to have 4.

\subsection{Missed Opportunites}
```{r overdemand, include=TRUE}
##### Over Demand Vector ####
overdemand <- 
  c(
    1 - sum(sapply(c(0:1), p)),
    1 - sum(sapply(c(0:2), p)),
    1 - sum(sapply(c(0:3), p)),
    1 - sum(sapply(c(0:4), p)),
    1 - sum(sapply(c(0:4), p))
    )

overdemand <- matrix(overdemand,
                     nrow = 1,
                     dimnames = list(
                       "", 
                       c("1 Tractor", paste(c(2:5), "Tractors"))
                     ))

##### Display Overdemand Nicely ####
kable(round(overdemand * 100, 3), caption = "Overdemand Probabilities")

##### What Percent of sales did we miss? ####
missed_sales <- sum(steady_state * overdemand)

```

Based on Acme's inventory practices, our calculated steady state, and the 
Poisson probability given $\lambda = 1$ we find that they will miss out on sales
approximately `r round(missed_sales * 100, 3)`\% of the time.
This number does not strike us as particularly egregious, however we leave it to 
the individuals with access to accounting data to determine whether this is 
worth investing in expanded capacity or if inventory practices need reexamining.

\section{Sensitivity}
We consider how four other sales assumptions would impact our model in the hopes
that we will arrive at a better understanding of the variety of 
possible outcomes.
```{r sensitivity_analysis, include=TRUE}
##### Model Sensitivity #####
# How sensitive is our model to our initial assumptions
i <- 1
sensitivity_matrix <- matrix(
  c(1, round(missed_sales * 100, 3)),
    nrow = 5, 
    ncol = 2,
   byrow = TRUE
  )
colnames(sensitivity_matrix) <- c("Weekly Sales", "Expected Missed Sales")

##### Evaluate Sensitivity ####
for (lambda in c(0.50, 0.75, 1.25, 1.50)) {
  ##### Sensitivity: Transition Matrix####
  transition_matrix <- matrix(
  c(p(0), 0, 0, 0, 1 - p(0),
    p(1), p(0), 0, 0, 1 - sum(sapply(c(0:1), p)),
    p(2), p(1), p(0), 0, 1 - sum(sapply(c(0:2), p)),
    p(3), p(2), p(1), p(0), 1 - sum(sapply(c(0:3), p)),
    p(4), p(3), p(2), p(1), 1 - sum(sapply(c(1:4), p))
    ),
  byrow = TRUE,
  nrow = 5,
  ncol = 5)
  
  ##### Sensitivity: Initial State ####
  initial_state <- transition_matrix[5, ]

  ##### Sensitivity: Steady State ####
  steady_state <- Freds_MarkovChain_SteadyState(initial_state, transition_matrix, -10, 100)

  ##### Sensitivity: Over Demand ####
  overdemand <- 
  c(
    1 - sum(sapply(c(0:1), p)),
    1 - sum(sapply(c(0:2), p)),
    1 - sum(sapply(c(0:3), p)),
    1 - sum(sapply(c(0:4), p)),
    1 - sum(sapply(c(0:4), p))
    )
  
  ##### Sensitivity: Missed Sales ####
  missed_sales <- sum(steady_state * overdemand)
  
  sensitivity_matrix[i+1, 1] <- lambda
  sensitivity_matrix[i+1, 2] <- round(missed_sales * 100, 3)
  i = i + 1
}

##### Display Sensitivity ####
kable(sensitivity_matrix, caption = "Sales Expectations")
```

Our table of sales expectations shows relatively large swings in missed sales.
We see that our smallest assumption, one sale every two weeks, translates to a 
reduction in missed sales by a factor of 3, while our largest assumption, 
three sales every two weeks, results in missing twice as many sales. 
The chart below has been provided for visual aide. 


```{r sensitivity_plot, include=TRUE}
##### Sensitivity Chart ####
sensitivity_matrix %>%
  as.data.frame() %>%
  ggplot(aes(x = `Weekly Sales`,
             y = `Expected Missed Sales`)) +
  geom_point() +
  geom_line() +
  ggtitle("Model Sensitivity to Variations in the Rate of Expected Sales")
```

\section{In Summation}
Had we more time
\footnote{As we write, deadlines and finals are fast approaching.}
we would have preferred to evaluate other aspects of our model than just
customer frequency assumptions. 
We would have liked to evaluate the impact of either reducing inventory to four
or increasing to six. 
If capital is harder to come by, Acme may find it more profitable to reduce
inventory and turn to capital markets to loan the remaining funds.
Regardless, we expect Acme's inventory restrictions to have a far greater 
impact on our model than assumed patronage, and generally find it to be a more
interesting question.

\section{And One More Thing}
Ultimately we would be remiss if we did not take the opportunity in our final 
submitted paper of the year to thank Dr. Ray Mugno for his tolerance and 
patience while being forced to read the many ERSA:$tz$ projects this semester.
Thanks Mugno. 

\clearpage
# Appendix I
## The Definition of Some `FredsStatsPack` Functions Used in this Analysis
```{r freds_markov, include=TRUE, echo=TRUE, eval=FALSE}
#' @title Freds_MarkovChain
#'
#' @param intial_state 1 X n matrix
#' @param transition_matrix a square n X n matrix 
#' @param iterations 
#'
#' @return 1 X n matrix indicating the current state
#'
#' @examples
Freds_MarkovChain <- function(initial_state, transition_matrix, iterations){
  
  ##### Need expm Package ####
  # For the matrix exponential operator %^%
  library(expm)
  
  ##### Matrix Exponentiate Transition Matrix ####
  transition_matrix <- transition_matrix %^% iterations
    
  ##### Do Math ####
  # Current state = initial state * transition matrix ^ iteration
  current_state <- initial_state %*% transition_matrix
  
  ##### Return state after iterations #### 
  return(current_state)
}
```

\clearpage

```{r freds_steady_state, include=TRUE, echo=TRUE, eval=FALSE}
#' @title Freds_MarkovChain_Steadystate
#'
#' @param initial_state       1 X n matrix indicating initial position
#' @param transition_matrix   n X n matrix 
#' @param tolerance           error bound to stop iterations, currently set to 
#'                              10 ^ -3 
#' @param max_iterations      maximum number of iterations to search for steady
#'                              state, currently set to 15
#'
#' @return A steady state if one exists within constraints
#'
Freds_MarkovChain_SteadyState <- function(initial_state, transition_matrix, 
                                         tolerance = -3,
                                    max_iterations = 15) {

  ##### Loop through the max iterations minus one ####
  for (iteration in 1:(max_iterations-1)) {
    
    ##### Difference between the states ####
    previous_state <- Freds_MarkovChain(initial_state, transition_matrix, iteration)
    
    current_state <- Freds_MarkovChain(initial_state, transition_matrix, iteration+1) 
    
    # Create boolean vector
    diff <- abs(current_state - previous_state) < 10 ^ tolerance
    
    ##### Stop Condiditions ####
    # If Steady State reached, return it 
    if (sum(diff) == length(diff)) {
      return(current_state)
      
      # If we haven't reached the Steady State by the max iterations, 
      # throw error saying so
    } else if (iteration == max_iterations-1) {
      stop("Steady State not reached in ", max_iterations, " iterations.")
    }
  }
  
}
```




\clearpage
# Appendix II
ERSA:$tz$ code used in the above analysis
```{r code_appendix, include=TRUE, echo=TRUE, eval=FALSE}
##### Chunk Options ####
opts_chunk$set(
     echo = FALSE,
  include = FALSE,
  message = FALSE,
  warning = FALSE,
  # Figure Settings 
  fig.asp = 0.618,
fig.width = 6)

##### Packages ####
library(FredsStatsPack)  # Our Very Own
library(magrittr)
library(data.table)
library(ggplot2)

##### Initial Conditions ####
# Poisson Probability
# We know lambda = 1
lambda <- 1
p <- function(k){lambda^k * exp(-lambda)/factorial(k)}

# Transition matrix
transition_matrix <- matrix(
  c(p(0), 0, 0, 0, 1 - p(0),
    p(1), p(0), 0, 0, 1 - sum(sapply(c(0:1), p)),
    p(2), p(1), p(0), 0, 1 - sum(sapply(c(0:2), p)),
    p(3), p(2), p(1), p(0), 1 - sum(sapply(c(0:3), p)),
    p(4), p(3), p(2), p(1), 1 - sum(sapply(c(1:4), p))
    ),
  byrow = TRUE,
  nrow = 5,
  ncol = 5)

# We take as the initial condition the last row
#   of the transition matrix
initial_state <- transition_matrix[5, ]

# Display transition matrix
kable(transition_matrix, caption = "Transition Matrix")

##### Markov Chain Matrix ####
state_mat <- matrix(nrow = 11, ncol = 5,
                    dimnames = list(
                      c("Initial Week", paste("Week", c(1:10))),
                      c("1 Tractor", paste(c(2:5), "Tractors"))))

# Start with initial condition 
# Fill matrix with weekly position
for (week in c(0:10)) {
  state_mat[week+1, ] <- round(Freds_MarkovChain(initial_state, transition_matrix, week) * 100, 3)
}

# x-axis for our plot
week <- data.table(c(0:10))
colnames(week) <- "Week"

##### Inventory Projection Table ####
kable(state_mat, caption = "Inventory Probability Matrix")

##### Plot Markov Chain ####
state_mat %>%
  cbind(week) %>%
  as.data.table() %>%
  melt.data.table(id.vars = "Week") %>%
  ggplot(aes(x = Week,
             y = value,
             color = variable)) +
  geom_line() +
  xlab("Week") +
  ylab("Probability of k Tractors Week Start") + 
  ggtitle("Tractor Inventory Projections")

##### Determine Steady State ####
# Find the steady state for tolerance 10e-17
steady_state <- Freds_MarkovChain_SteadyState(initial_state, transition_matrix, -17, 100)

steady_state <- matrix(steady_state,
                       nrow = 1,
                       dimnames = list(
                         "",
                         c("1 Tractor", paste(c(2:5), "Tractors"))
                       ))

##### Format Steady State nicely #####
kable(steady_state, caption = "Tractor Steady State")

##### Over Demand Vector ####
overdemand <- 
  c(
    1 - sum(sapply(c(0:1), p)),
    1 - sum(sapply(c(0:2), p)),
    1 - sum(sapply(c(0:3), p)),
    1 - sum(sapply(c(0:4), p)),
    1 - sum(sapply(c(0:4), p))
    )

overdemand <- matrix(overdemand,
                     nrow = 1,
                     dimnames = list(
                       "", 
                       c("1 Tractor", paste(c(2:5), "Tractors"))
                     ))

##### Display Overdemand Nicely ####
kable(overdemand, caption = "Overdemand Probabilities")

##### What Percent of sales did we miss? ####
missed_sales <- sum(steady_state * overdemand)
##### Model Sensitivity #####
# How sensitive is our model to our initial assumptions
i <- 1
sensitivity_matrix <- matrix(
  c(1, round(missed_sales * 100, 3)),
    nrow = 5, 
    ncol = 2,
   byrow = TRUE
  )
colnames(sensitivity_matrix) <- c("Weekly Sales", "Expected Missed Sales")

##### Evaluate Sensitivity ####
for (lambda in c(0.50, 0.75, 1.25, 1.50)) {
  ##### Sensitivity: Transition Matrix####
  transition_matrix <- matrix(
  c(p(0), 0, 0, 0, 1 - p(0),
    p(1), p(0), 0, 0, 1 - sum(sapply(c(0:1), p)),
    p(2), p(1), p(0), 0, 1 - sum(sapply(c(0:2), p)),
    p(3), p(2), p(1), p(0), 1 - sum(sapply(c(0:3), p)),
    p(4), p(3), p(2), p(1), 1 - sum(sapply(c(1:4), p))
    ),
  byrow = TRUE,
  nrow = 5,
  ncol = 5)
  
  ##### Sensitivity: Initial State ####
  initial_state <- transition_matrix[5, ]

  ##### Sensitivity: Steady State ####
  steady_state <- Freds_MarkovChain_SteadyState(initial_state, transition_matrix, -10, 100)

  ##### Sensitivity: Over Demand ####
  overdemand <- 
  c(
    1 - sum(sapply(c(0:1), p)),
    1 - sum(sapply(c(0:2), p)),
    1 - sum(sapply(c(0:3), p)),
    1 - sum(sapply(c(0:4), p)),
    1 - sum(sapply(c(0:4), p))
    )
  
  ##### Sensitivity: Missed Sales ####
  missed_sales <- sum(steady_state * overdemand)
  
  sensitivity_matrix[i+1, 1] <- lambda
  sensitivity_matrix[i+1, 2] <- round(missed_sales * 100, 3)
  i = i + 1
}

##### Display Sensitivity ####
kable(sensitivity_matrix, caption = "Sales Expectations")

##### Sensitivity Chart ####
sensitivity_matrix %>%
  as.data.frame() %>%
  ggplot(aes(x = `Weekly Sales`,
             y = `Expected Missed Sales`)) +
  geom_point() +
  geom_line() +
  ggtitle("Model Sensitivity to Variations in the Rate of Expected Sales")

```



